maxent model avec xvalidation (year-based test training)

### 2019-05-07 src/maxent_avec_xvalidation.Rmd
missing loop for betaval back in

select month
```{r month of interest -change}
mthna <- "mar"
mthno <- "3"
```

@@

  for (h in 1:nrow(betavallist)){
    betamval <-  betavallist[h]
    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))

@@

      aicetc$train_AUC <- trainauc
      aicetc$no_trainng_points <- notrainingpoints
      aicetc$test_coefficients <- nocoeff
      aicetc$test_sum.of.logs <- sumsamplepred
      aicetc$test_AUC <- evauc
      aicetc$test_evcor <- evcor
      aicetc$test_evpcor <- evpcor
      aicetc$test_nopresence <- evnp
      aicetc$test_nobackground <- evna
      aicetc$test_maxsumsenspec <- evmaxtprtnr
      aicetc$test_maxkap <- evmaxkap
      aicetc$test_no_omission <- evno_omission
      aicetc$test_prevalence <- evprevalence
      aicetc$test_equal_sens_spec <- evequal_sens_spec
      aicetc$test_sensitivity <- evsensitivity
      aicetc$test_bic <- bic #not sure this is correct...
      aicetc$beta <- betaval

    #calculate minimal predicted area using ecospat
    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
      }
    }

@@

@@

```{r final models plus predict}
modelmonth <- cbind(paste0(mthna, unique(prback$year)))
aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
unlayer <- read.csv(paste0("../data/env/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr

@@

  if (yr == "2003"){
    betaval <- 1
  } else if (yr == "1999" | yr == "2002" | yr == "2008" | yr == "2009") {
    betaval <- 1.25
  } else if (yr == "2005" | yr == "2006"){
    betaval <- 1.5
  } else if (yr == "2000"){
    betaval <- 1.75
  } else if (yr == "2001" | yr == "2007"){
    betaval <- 2.25
  } else if (yr == "2010"){
    betaval <- 3.5
  }

@@

```{r binary map per year 2D representation}
binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_.*.asc", full.names = TRUE)
yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryclog_list)))
colnames(yrlayer) <- c("layer", "year")

for (binyr in (1:length(y_list))){
  binstack <- stack(y_list[[binyr]]$layer)
  year <- y_list[[binyr]]$year[1]
  stacksum <- sum(binstack, na.rm = TRUE)
  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".asc"))
  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".png")))
  dev.off()
}

@@

aic <- rbind(aic2003, aic2004, aic2005, aic2006)
maxr <- rbind(maxr2003, maxr2004, maxr2005, maxr2006)

@@

@@

```{r avg asc layers cloglog}
clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
stdep <- read.csv(paste0("../output/maxent/predictions/", mthna, "/stack_depth.csv"), header = TRUE)
depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
depthlayers$depth <- as.character(depthlayers$depth)

@@

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1)
  asclay$V2 <- as.character(asclay$V2)
  asclay$V3 <- as.character(asclay$V3)
  asclay$V4 <- as.character(asclay$V4)
  asclay$V5 <- as.character(asclay$V5)
  asclaydep <- asclay$V1

  asc2003 <- raster(asclay$V2)
  asc2004 <- raster(asclay$V3)
  asc2005 <- raster(asclay$V4)
  asc2006 <- raster(asclay$V5)

  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
  avg <- mean(stkasc, na.rm = TRUE)

@@

binstack <- stack(binaryclog_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.asc"))
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.png")))
dev.off()

@@

```{r graph by depth (avg)}
bydepth <- barchart(layer ~ cell_sum, data = stackstat, ylim = c(49, 1), xlim = c(0, 300))
trellis.device(device="png", filename = paste0("../output/maxent/predictions/", mthna, "/depthplot_avg.png"))
print(bydepth)
dev.off()
```

>>>>

select month
```{r month of interest -change}
mthna <- "may"
mthno <- "5"
```

@@

  for (h in 1:nrow(betavallist)){
    betaval <-  betavallist[h]
    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))

@@

    for (i in 1:nrow(aicetc)){
    if(aicetc$regularization[i] == betaval){
      aicetc$train_AUC[i] <- trainauc
      aicetc$no_trainng_points[i] <- notrainingpoints
      aicetc$test_coefficients[i] <- nocoeff
      aicetc$test_sum.of.logs[i] <- sumsamplepred
      aicetc$test_AUC[i] <- evauc
      aicetc$test_evcor[i] <- evcor
      aicetc$test_evpcor[i] <- evpcor
      aicetc$test_nopresence[i] <- evnp
      aicetc$test_nobackground[i] <- evna
      aicetc$test_maxsumsenspec[i] <- evmaxtprtnr
      aicetc$test_maxkap[i] <- evmaxkap
      aicetc$test_no_omission[i] <- evno_omission
      aicetc$test_prevalence[i] <- evprevalence
      aicetc$test_equal_sens_spec[i] <- evequal_sens_spec
      aicetc$test_sensitivity[i] <- evsensitivity
      aicetc$test_bic[i] <- bic #not sure this is correct...
      aicetc$beta[i] <- betaval
      #calculate minimal predicted area using ecospat
      aicetc$mpa_0.9_cum[i] <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%

    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
    }
   }
  }

@@

you need to go through each aic file manually....

@@

#run final model
```{r final models plus predict}
modelmonth <- cbind(paste0(mthna, unique(prback$year)))
aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
unlayer <- read.csv(paste0("../data/env/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr

@@

  if (yr == "2001" | yr == "2006" | yr == "2007"){
    betaval <- 1.75
  } else if (yr == "2000"| yr == "2002" | yr == "2003" | yr == "2004"){
    betaval <- 2
  } else if (yr == "1999" | yr == "2005" | yr == "2008" | yr == "2009"){
    betaval <- 2.25
  } else if (yr == "2010" | yr == "2011"){
    betaval <- 2.5
  } else if (yr == "1998"){
    betaval <- 2.75
  }

@@

@@

aic <- rbind(aic1998, aic1999, aic2000, aic2001, aic2002, aic2003, aic2004, aic2005, aic2006, aic2007, aic2008, aic2009, aic2010, aic2011)
maxr <- rbind(maxr1998, maxr1999, maxr2000, maxr2001, maxr2002, maxr2003, maxr2004, maxr2005, maxr2006, maxr2007, maxr2008, maxr2009, maxr2010, maxr2011)

@@

```{r binary map per year 2D representation}
binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_cloglog.*.asc", full.names = TRUE)
yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryclog_list)))
colnames(yrlayer) <- c("layer", "year")

for (binyr in (1:length(y_list))){
  binstack <- stack(y_list[[binyr]]$layer)
  year <- y_list[[binyr]]$year[1]
  stacksum <- sum(binstack, na.rm = TRUE)
  writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".asc"), overwrite=TRUE)
  plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_2d_", year, ".png")))
  dev.off()
}

@@

```{r avg asc layers cloglog}
clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^cloglog_.*.asc", full.names = TRUE)
stdep <- read.csv(paste0("../output/maxent/models/", timeslice, "/stack_depth.csv"), header = TRUE)
depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
depthlayers$depth <- as.character(depthlayers$depth)

@@

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
  asclay$V3 <- as.character(asclay$V3)
  asclay$V4 <- as.character(asclay$V4)
  asclay$V5 <- as.character(asclay$V5)
  asclay$V6 <- as.character(asclay$V6)
  asclay$V7 <- as.character(asclay$V7)
  asclay$V8 <- as.character(asclay$V8)
  asclay$V9 <- as.character(asclay$V9)
  asclay$V10 <- as.character(asclay$V10)
  asclay$V11 <- as.character(asclay$V11)
  asclay$V12 <- as.character(asclay$V12)
  asclay$V13 <- as.character(asclay$V13)
  asclay$V14 <- as.character(asclay$V14)
  asclay$V15 <- as.character(asclay$V15)

  asclaydep <- asclay$V1

  asc1998 <- raster(asclay$V2)
  asc1999 <- raster(asclay$V3)
  asc2000 <- raster(asclay$V4)
  asc2001 <- raster(asclay$V5)
  asc2002 <- raster(asclay$V6)
  asc2003 <- raster(asclay$V7)
  asc2004 <- raster(asclay$V8)
  asc2005 <- raster(asclay$V9)
  asc2006 <- raster(asclay$V10)
  asc2007 <- raster(asclay$V11)
  asc2008 <- raster(asclay$V12)
  asc2009 <- raster(asclay$V13)
  asc2010 <- raster(asclay$V14)
  asc2011 <- raster(asclay$V15)

  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011)

  avg <- mean(stkasc, na.rm = TRUE)

@@

binstack <- stack(binaryclog_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.asc"), overwrite=TRUE)
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_avg_2d.png")))
dev.off()

@@

```{r graph by depth (avg)}
bydepth <- barchart(layer ~ cell_sum, data = stackstat, ylim = c(49, 1))
trellis.device(device="png", filename = paste0("../output/maxent/predictions/", mthna, "/depthplot_avg.png"))
print(bydepth)
dev.off()
```
